---
title: "Untitled"
author: "April Truong"
date: "February 16, 2019"
output: 
  html_document: 
    keep_md: yes
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Online you will find the filestudent.csv. It containsinformation on 1548 introduction to statistics students,and has the following columns:
Column 1:height:  The height of the student.  
Column 2:hsGPA: The high school GPA of the student.  
Column 3:pulse:  The pulse rate of the student whenmeasured in class.  

#read data
```{r}
student <- readr::read_csv("C:/Users/Apple/Downloads/R Files/student.csv")
library(tidyverse)
```

#(a) Plot a QQ plot (normal probability plot) of students heights
```{r}
colnames(student)

qqnorm(student$height, main = "Normal Probability Plot for Student Height")
qqline(student$height)
```
> yes  

#(b) Plot a QQ plot (normal probability plot) for students pulse.
```{r}
qqnorm(student$pulse,main = "Normal Probability Plot for Student Pulse")
qqline(student$pulse)
```
> no

#(c) Plot a QQ plot (normal probability plot) for highschool GPA.
```{r}
qqnorm(student$hsGPA,main = "Normal Probability Plot for Student Hs GPA")
qqline(student$hsGPA)
```
>no

#(d) What proportion of the data for students height lies within one standard deviation from the mean?
```{r}
X = student$height
the.mean = mean(X)
the.sd = sd(X)
one.sd.away = mean(X < the.mean + the.sd & X > the.mean - the.sd)
one.sd.away
```
>73.5%

#(e) What proportion of the data for students pulse lies within two standard deviation from the mean?
```{r}
X = student$pulse
the.mean = mean(X)
the.sd = sd(X)
two.sd.away = mean(X < the.mean + 2*the.sd & X > the.mean - 2*the.sd)
two.sd.away
```
>96%

#(f) What proportion of the data for highschool GPA lies within three standard deviation from the mean?
```{r}
X = student$hsGPA
the.mean = mean(X)
the.sd = sd(X)
three.sd.away = mean(X < the.mean + 3*the.sd & X > the.mean - 3*the.sd)
three.sd.away 
```
>0.985

