---
title: "HW#1 (pain)"
author: "April Truong"
date: "February 16, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#read data
```{r}
pain <- read.csv("C:/Users/Apple/Downloads/R Files/pain.csv")
```

# On Canvas, you will find the dataset"pain.txt". This  dataset has the following columns:
Column 1:  HairColor:  The hair color of the subject
Column 2:  Score:  The pain tolerance of the subject
Each person in the experiment was given a pain thresh-old score based on his or her performance in a pain sensitivity test (the higher the score, the higher the person's pain tolerance).

##(a)  Find the overall mean of pain tolerance, regardlessof hair color.
```{r}
colnames(pain)
head(pain)
mean(pain$Score)
```

##(b)  Find  the  overall  standard  deviation  of  pain  toler-ance, regardless of hair color.
```{r}
sd(pain$Score)
```

##(c)  Find the mean pain tolerance for the category DarkBrunette. Hint: Use R to find the mean for each group.  
```{r}
aggregate( Score ~ HairColor, data = pain, mean)
```

##(d)  Find the  standard deviation of pain tolerance for the category LightBrunette. Hint: Use R to find the standard deviation for each group.  
```{r}
aggregate( Score ~ HairColor, data = pain, sd)
```

##(e)  Find the number of subjects that were in the cate-gory LightBlond.Hint:  Use R to find the number of subjects for each group.  
```{r}
aggregate( Score ~ HairColor, data = pain, length)
```

##(f)  Which hair color category had the highest average pain tolerance?  
```{r}

```

#(g)  Which hair color category had the lowest standarddeviation of pain tolerance?  
```{r}

```

#(h)  Which hair color category had the most subjects?
```{r}

```

